Global Value Chain Analysis of LMICs

### 2016-05-19 Draft.tex
final version

\usepackage{mathtools}
\usepackage{threeparttable}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{changepage}
\usepackage{verbatim}
\usepackage{epigraph}
\usepackage{lscape}
\usepackage{pbox}
\usepackage{array}
\usepackage{natbib}
\usepackage{graphicx}

@@

A central step towards a more in-depth analysis of GVCs has been laid by \citet{rokoetal14} and \citet{zhwaetal13} who show that it is necessary to go beyond deriving origins of value added to examine production sharing comprehensively. They split exports into different categories and calculate metrics of how often goods cross borders. This enables them to derive measures of GVC length but also allows to investigate how individual countries are integrated into GVCs. For instance, they show that a considerable part of US value added exports eventually returns home in the form of final goods which is indicative of the US being specialised in upstream production stages.\footnote{See \citeauthor{joamsoca15} (forthcoming) for a comprehensive review of the literature on GVCs and outsourcing.}

However, these contributions typically have two shortcomings. Firstly, most evidence is based on data from high-income countries. The reason is that reliable time-series of both national and international input-output tables have only been available for this particular subset of countries. In addition, the evidence is regularly based on a small sample of GVC indicators that hide valuable information stemming from more decomposed and disaggregated indicators.

@@

Our analysis confirms the expansion of GVCs in recent years and presents evidence that GVCs have become longer over time. We also find that these developments are increasingly driven by low- and middle-income countries while the integration of high-income countries has begun to even out at a high level. In addition, we find that high-income countries typically are the starting and end points of GVCs in that they provide upstream inputs and then serve eventually again as demand markets for the final products. Low- and middle-income countries, on the other hand, are rather specialised in downstream activities such as assembly and export typically less domestic value added. However, we also observe that developing economies move more and more upstream along the value chain and should thus be able to generate more gains from their GVC participation.

@@

The idea is that the production of industry \textit{i} creates value added in industry $i$ itself, a direct contribution, but also in industries that supply $i$ with inputs, an indirect contribution. Since these industries themselves rely on inputs, $i$'s production sets several rounds of indirect value added creation in motion that can mathematically be expressed as:
\begin{equation}
VB = V + VA + VAA + VAAA + ... = V (I+A+A^{2}+A^{3}+...),
\end{equation}
which, as an infinite geometric series with the elements of $A<1$, simplifies to
\begin{equation}
VB = V (I-A)^{-1},
\end{equation}
where \textit{V} is a matrix with the diagonal representing the direct value added contribution of each industry, \textit{A} is the Input-Output coefficient matrix, which means it gives the direct input flows between industries required for 1\$ of output, and $B = (I-A)^{-1}$ is the so called Leontief inverse. \textit{VB} gives thus so called value added multipliers, which denote the amount of value added that the production of an industry's 1\$ of output or exports brings about in all other industries. If we post-multiply $VB$ with exports, we get a matrix with the value added origins of each industry's exports.

This basic decomposition has been widely used in GVC analysis since it allows the calculation of two informative GVC participation measures, namely foreign value added in exports (Hummels et al.'s Vertical Specialisation) and domestic value added in foreign exports (Hummels et al.'s Vertical Specialisation 1). These indicators can tell us how much a country is integrated into GVCs and if it acts mainly as a supplier or a user of foreign value added. However, the Leontief decomposition is only informative for the origin and destination of value added while ICIOs also contain info on the type of good that is being traded and and how often an intermediate crosses borders. The WWZ decomposition extends the Leontief decomposition in this direction and thereby extracts more insights from ICIOs.

@@

Since the derivation itself is not the focus of this paper, we present here only the final result for a $G$-country $N$-industry model (equation 37 in WWZ) and refer the interested reader to the original paper. WWZ use the Leontief decomposition and extend it using additional information from ICIOs on the final usage and destination of the exports (e.g. re-imported vs. absorbed abroad). This splits the exports, $E$, of industry $l$ in country $k$ into sixteen different parts broadly differentiated into the four boroad categories domestic value added absorbed abroad, domestic value added returning home, foreign value added, and purely double counted terms:
\begin{align}\label{eq:wwz}
\begin{split}
E^{kl}
= &\left(V^k B^{kk} \right)^T * F^{kl} 
+ \left(V^k L^{kk} \right)^T * \left(A^{kl} B^{ll} F^{ll} \right)\\
+& \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k,l}^G  B^{lt} F^{tt} )
+ \left(V^k L^{kk} \right)^T *  (A^{kl} B^{ll} \sum_{t \neq k,l}^G  F^{lt} )\\ 
+&  \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k}^G \sum_{l,u \neq k,t}^G B^{lt} F^{tu} )
+ \left(V^k L^{kk} \right)^T * \left(A^{kl} B^{ll} F^{lk} \right)\\
+& \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k,l}^G  B^{lt} F^{tk} )
+ \left(V^k L^{kk} \right)^T * \left(A^{kl} B^{lk} F^{kk} \right) \\
+& \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k}^G  B^{lk} F^{kt} )
+ \left(V^k B^{kk} -  V^k L^{kk} \right)^T * \left(A^{kl} X^{l}  \right)\\
+& \left(V^l B^{lk} \right)^T * F^{kl}
+ \left(V^l B^{lk} \right)^T *  \left(A^{kl} L^{ll} F^{ll} \right)
+ \left(V^l B^{lk} \right)^T \\
*&  \left(A^{kl} L^{ll} E^{l*} \right) + (\sum_{t \neq k,l}^G  V^{t} B^{tk} )^{T} * F^{kl}
+ (\sum_{t \neq k,l}^G  V^{t} B^{tk} )^{T}\\
*&   \left(A^{kl} L^{ll} F^{ll} \right) + (\sum_{t \neq k,l}^G  V^{t} B^{tk} )^{T} *  \left(A^{kl} L^{ll} E^{l*} \right) ,
\end{split}
\end{align}
where $F$ is final demand, and $L$ refers to the domestic Leontief inverse as opposed to the global inverse $B$. $X$ is output while \textit{T} indicates a matrix transpose operation.

@@

\begin{itemize}
\item Domestic value added absorbed abroad (\textit{VAX\_G}, T1-5)
\begin{itemize}
\item Domestic value added in final exports (\textit{DVA\_FIN}, T1)
\item Domestic value added in intermediate exports
\begin{itemize}
\item Domestic value added in intermediate exports absorbed by direct importers (\textit{DVA\_INT}, T2)
\item Domestic value added in intermediate exports re-exported to third countries (\textit{DVA\_INTrex}, T3-5)
\begin{itemize}
\item Domestic value added in intermediate exports re-exported to third countries as intermediate goods to produce domestic final goods (\textit{DVA\_INTrexI1}, T3)
\item Domestic value added in intermediate exports re-exported to third countries as  final goods (\textit{DVA\_INTrexF}, T4)
\item Domestic value added in intermediate exports re-exported to third countries as intermediate goods to produce exports (\textit{DVA\_INTrexI2}, T5)
\end{itemize}
\end{itemize}
\end{itemize}
\item Domestic value added returning home (\textit{RDV\_B}, T6-8)
\begin{itemize}
\item Domestic value added returning home as final goods (\textit{RDV\_FIN}, T6)
\item Domestic value added returning home as final goods through third countries (\textit{RDV\_FIN2}, T7)
\item Domestic value added returning home as intermediate goods (\textit{RDV\_INT}, T8)
\end{itemize}
\item Foreign value added (\textit{FVA}, T11-12/14-15 )
\begin{itemize}
\item Foreign value added in final good exports (\textit{FVA\_FIN}, T11/14)
\begin{itemize}
\item Foreign value added in final good exports sourced from direct importer (\textit{MVA\_FIN}, T11)
\item Foreign value added in final good exports sourced from other countries (\textit{OVA\_FIN}, T14)
\end{itemize}
\item Foreign value added in intermediate good exports (\textit{FVA\_INT}, T12/15)
\begin{itemize}
\item Foreign value added in intermediate good exports sourced from direct importer (\textit{MVA\_INT}, T12)
\item Foreign value added in intermediate good exports sourced from other countries(\textit{OVA\_INT}, T15)
\end{itemize}
\end{itemize}
\item Pure double counting (\textit{PDC}, T9-10/13/16)
\begin{itemize}
\item Pure double counting from domestic source (\textit{DDC}, T9-10)
\begin{itemize}
\item Due to final goods exports production (\textit{DDF}, T9)
\item Due to intermediate goods exports production (\textit{DDI}, T10)
\end{itemize}
\item Pure double counting from foreign source (\textit{FDC}, T13/16)
\begin{itemize}
\item Due to direct importer exports production (\textit{FDF}, T13)
\item Due to other countries' exports production (\textit{FDI}, T16)
\end{itemize}
\end{itemize}
\end{itemize}

The advantage of such a detailed decomposition is that these new indicators can inform us on how countries integrate into GVCs while the basic Leontief decomposition mainly informs us on the intensity of integration. High amounts of foreign value added in final goods exports are, for instance, suggestive of a specialisation in downstream tasks that add little value to a good, such as assembly. High amounts of domestic value added in intermediate exports, on the other hand, are evidence of a more upstream specialisation in tasks that aaa a lot of value, such as business services. By tracking these two variables over time we can then see which countries have succeeded in moving up the value chain.

The high resolution of the WWZ decomposition does, however, not mean that the Leontief decomposition does not contain valuable information at all. In fact, we exploit the decomposition of exports into source industry and source country by calculating variants of the standard indicators based on different characteristics. In particular, we will assess the integration of low- and middle-income countries into GVCs by computing the amount of value added that they supply for total GVC trade.

@@

In this section we look with our extensive OECD ICIO dataset into stylised facts on GVC integration that are typically based on smaller samples. We start by examining the development of the most basic measure of GVC integration introduced by \citet{dahuetal01}, namely the amount of foreign value added in exports labeled by \citet{ribajalo13} as $i2e$.\footnote{Note that at the aggregate level forward and backward linkages are identical and thus we only look at one of the two measures.} It captures backward linkages into value chains and shows the well-known increase in GVC integration from 1995 to 2011. As illustrated in Figure \ref{fig:i2e_t}, the value of $i2e$ has grown by approximately 350\% and by 35\% as a share of total exports from around 17\% to over 23\%. Thus, countries rely for their export production increasingly on inputs produced abroad. The numbers are in line with similar findings by \citet{rojoguno12b} but their sample ends in 2009. It is then interesting to see that after the slump during the financial crises in 2009, GVCs have quickly recovered and already have started to exceed their pre-crisis levels in 2011.

%\begin{figure}[h]
%\centering
%\label{fig:i2e_t}
%\includegraphics[width=7cm]{i2epercentage.png}
%\caption{The development of GVC integration over time}
%\end{figure}

Another way to examine the expansion of GVCs from 1995 to 2011 is to look at their intensity or length instead of their trade volume. \citet{zhwaetal13} propose to use the amount of double counted trade, $pdc$, as a proxy for GVC length since its value goes up the more back-and-forth trade occurs, which is equivalent to an increase in the number of production stages. They show that its value has increased for the 40 WIOD countries. In Figure \ref{fig:pdc_t}, we observe similarly that $pdc$ as a share of total exports has increased over the examined period by 73\% and thus more than $i2e$. Therefore, GVCs do not only channel more trade but also have become longer over time.

@@

Turning from the development over time to sectoral differences in GVC integration, Figure \ref{fig:i2e_s} reveals in line with \citet{rojoguno12a} that the sectors exhibiting the highest dependency on foreign inputs in terms of $i2e$ shares are heavy manufactures such as motor vehicles (MTR), other transport equipment (TRQ) and the metal industry (MET) as well as computers and electronics (CEQ and ELQ). In particular, the transport equipment and electronics industry are strongly engaged in GVCs having highly international production networks. For instance, Apple's iPhone contains inputs from 9 to 10 countries while the Boeing 787 production spans more than 5 countries. The sectors can be characterised as being close to final demand or being dependent on imported raw materials. These characteristics can thus explain differences in GVC integration.

The bottom 6 industries in terms of $i2e$ shares are primary and services sectors such as agriculture (AGR), mining (MIN), R\&D and business services (BZS), or wholesale and retail trade (WRT). These sectors are typically located upstream in the supply chain far from final demand and have high high value added to output ratios.

\begin{figure}
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{industries_i2e_top6.png}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{industries_i2e_bottom6.png}
\end{minipage}
\caption{Sectoral backward linkages into GVCs - Top and bottom 6.}
\label{fig:i2e_s}
\end{figure}

Naturally then, things are reversed when we look at the corresponding forward linkage GVC measure, $e2r$. It captures the amount of domestic value added in foreign exports and thus quantifies how important domestic industries are for foreign export production. Here, Figure \ref{fig:e2r_s} depicts that this indicator is dominated by the same upstream industries that are at the bottom of the $i2e$ ranking such as mining or financial and telecommunication services (FIN and PTL). This shows that these industries are also strongly engaged in GVCs but their participation is of a different type. They primarily supply important inputs but do not serve final demand.

With respect to services, it is also indicative of the servicification of manufacturing as described by \citet{ribaetal15}. This means that an increasing share of manufacturing gross exports is actually value added generated in services sectors and then embedded in the intermediate exports of manufacturers. This importance of services sectors for exports cannot be seen from standard gross trade statistics and thus constitutes a major advantage of trade in value added measures. It is also indicative of a growing internationalisation of services. Increasingly, services are being offshored and sourced from abroad. In that respect, it is also interesting to note that despite the low absolute $i2e$ shares, it is in services where a lot of the growth in $i2e$ has taken place. Five out of the six sectors with the highest growth in $i2e$ shares are services sectors.

\begin{figure}
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{industries_e2r_top6.png}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{industries_e2r_bottom6.png}
\end{minipage}
\caption{Sectoral forward linkages into GVCs - Top and bottom 6.}
\label{fig:e2r_s}
\end{figure}

Finally, when we turn to differences in GVC integration by country we can confirm the findings by \citet{ribajalo13}. Figure \ref{i2e_k} shows that small countries close to the major GVC hubs in Asia, Europe, and North America such as Slovakia or Malaysia have the highest average $i2e$ shares. Countries specialised in the primary sector or assembly on the other hand have very low values. Correspondingly, Latin American countries with their focus on agriculture and mining have very weak backward linkages into GVCs. However, the development over time shows that some of the countries with the relatively low GVC integration have begun to catch up. For instance, Argentina, Indian and Turkey are in the top 6 when it comes to the growth of $i2e$ shares from 1995 to 2011.

\begin{figure}
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{country_i2e_top_6.png}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{countries_i2e_bottom_6.png}
\end{minipage}
\caption{Countries' backward linkages into GVCs - Top and bottom 6.}
\label{fig:i2e_k}
\end{figure}

Driven by the sectoral statistics, we then find again that for $e2r$ the picture is reversed with raw material exporters on top. If we abstract from these countries we find technologically advanced countries such as Switzerland and the main GVC hubs Japan, USA, and Germany to exhibit strong forward linkages into GVCs. In particular low and middle-income countries without raw materials such as Cambodia, Mexico, or Turkey have in contrast very weak linkages and have not been able to strengthen them significantly between 1995 and 2011.

\begin{figure}
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{country_e2r_top_6.png}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{countries_e2r_bottom_6.png}
\end{minipage}
\caption{Countries' forward linkages into GVCs - Top and bottom 6.}
\label{fig:e2r_k}
\end{figure}

@@

The central advantage of our application is that we have new indicators for a new set of countries. In particular, the OECD ICIO database extends the WIOD database by the following 21 regions: Argentina, Brunei Darussalam, Cambodia, Chile, Colombia, Costa Rica, Croatia, Hong Kong, Iceland, Israel, Malaysia, Norway, New Zealand, Philippines, Saudi Arabia, Singapore, Thailand, Tunisia, Vietnam, South Africa, and Switzerland. This means that in particular the coverage low and middle income countries has increased considerably which allows us to analyse the GVC integration in developing economies in a much more detailed fashion than before. In particular, we will merge a subset of the new countries into the three regions Central and South America (CSA), South East Asia (SEA), and Africa (AFR) and analyse the development of their GVC participation\footnote{CSA covers Argentina, Chile, Colombia, and Costa Rica; SEA covers Cambodia, Malaysia, Philippines, Thailand, and Vietnam; AFR covers South Africa and Tunisia.}.

\subsection{General trends in the participation of developing economies}

>>>>

\usepackage{mathtools}
\usepackage{threeparttable}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{changepage}
\usepackage{verbatim}
\usepackage{epigraph}
\usepackage{lscape}
\usepackage{pbox}
\usepackage{array}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{caption}

@@

A central step towards a more in-depth analysis of GVCs has been laid by \citet{rokoetal14} and \citet{zhwaetal13} who show that it is necessary to go beyond deriving origins of value added to examine production sharing comprehensively. They split goods into different categories and calculate metrics of how often these goods cross borders. This enables them to derive measures of GVC length but also allows to investigate how individual countries are integrated into GVCs. For instance, they show that a considerable part of US value added exports eventually returns home in the form of final goods which is indicative of the US being specialised in upstream production stages.\footnote{See \citeauthor{joamsoca15} (forthcoming) for a comprehensive review of the literature on GVCs and outsourcing.}

However, these contributions typically have one of two shortcomings. Firstly, most evidence is based on data from high-income countries. The reason is that reliable time-series of both national and international input-output tables have only been available for this particular subset of countries. In addition, the evidence is regularly based on a small sample of GVC indicators that hide valuable information stemming from more decomposed and disaggregated indicators.

@@

Our analysis confirms the expansion of GVCs in recent years and presents evidence that GVCs have become longer over time. We also find that these developments are increasingly driven by low- and middle-income countries while the integration of high-income countries has begun to even out at a high level. In addition, we find that high-income countries typically are the starting and end points of GVCs in that they provide upstream inputs and then serve eventually again as demand markets for the final products. Low- and middle-income countries, on the other hand, are more specialised in downstream activities such as assembly and export typically less domestic value added. However, we observe that developing economies have begun to move upstream along the value chain and out of pure assembly occupying a wider set stages. This should allow them to generate more gains from GVC participation.

@@

The idea is that the production of industry \textit{i} of country \textit{k} creates value added in industry $i$ itself, a direct contribution, but also in industries \textit{j} from \textit{k} or other countries \textit{l} that supply $i$ with inputs, an indirect contribution. Since these industries themselves rely on inputs, $i$'s production sets several rounds of indirect value added creation in motion that can mathematically be expressed as:
\begin{equation}
VB = V + VA + VAA + VAAA + ... = V (I+A+A^{2}+A^{3}+...),
\end{equation}
which, as an infinite geometric series with the elements of $A<1$, simplifies to
\begin{equation}
VB = V (I-A)^{-1},
\end{equation}
where \textit{V} is a matrix with the diagonal representing the direct value added contribution of each industry, \textit{A} is the Input-Output coefficient matrix, which means it gives the direct input flows between industries required for 1\$ of output, and $B = (I-A)^{-1}$ is the so called Leontief inverse. \textit{VB} gives thus so called value added multipliers, which denote the amount of value added that the production of an industry's 1\$ of output or exports brings about in all other industries. If we post-multiply $VB$ with exports, we get a matrix, $VAE$, with the elements being the value added origins of each industry's exports, $vae_{ikjl}$.

This basic decomposition has been widely used in GVC analysis since it allows the calculation of two informative GVC participation measures. Firstly, a backward linkage indicator that is given by the import content of exports, $i2e$, (\citet{dahuetal01}`s Vertical Specialisation) and calculated as follows:
\begin{equation}
i2e_{ik} = \frac{\sum_{l}{\sum_{j}vae_{jlik}}}{exports_{ik}} ,
\end{equation}
Secondly, a forward linkage indicator - \emph{e2r} (domestic content in foreign (re-)exports) - which is given by:
\begin{equation}\label{eq:dvar}
{e2r}_{ik} = \frac{\sum_{l}{\sum_{j} vae_{ikjl}}}{exports_{ik}} ,
\end{equation}
where \(l\  \neq k\).

These indicators can tell us how much a country is integrated into GVCs and if it acts mainly as a supplier or a user of foreign value added. However, the Leontief decomposition is only informative for the origin and destination of value added while ICIOs also contain info on the type of good that is being traded and and how often an intermediate crosses borders. The WWZ decomposition extends the Leontief decomposition in this direction and thereby extracts more insights from ICIOs.

@@

Since the derivation itself is not the focus of this paper, we present here only the final result for a $G$-country $N$-industry model (equation 37 in WWZ) and refer the interested reader to the original paper. WWZ use the Leontief decomposition and extend it using additional information from ICIOs on the final usage and destination of the exports (e.g. re-imported vs. absorbed abroad). This splits the exports, $E$, of industry $l$ in country $k$ into sixteen different parts broadly differentiated into the four broad categories domestic value added absorbed abroad, domestic value added returning home, foreign value added, and purely double counted terms:
\begin{align}\label{eq:wwz}
\begin{split}
E^{kl}
= &\left(V^k B^{kk} \right)^T * F^{kl} 
+ \left(V^k L^{kk} \right)^T * \left(A^{kl} B^{ll} F^{ll} \right)\\
+& \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k,l}^G  B^{lt} F^{tt} )
+ \left(V^k L^{kk} \right)^T *  (A^{kl} B^{ll} \sum_{t \neq k,l}^G  F^{lt} )\\ 
+&  \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k}^G \sum_{l,u \neq k,t}^G B^{lt} F^{tu} )
+ \left(V^k L^{kk} \right)^T * \left(A^{kl} B^{ll} F^{lk} \right)\\
+& \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k,l}^G  B^{lt} F^{tk} )
+ \left(V^k L^{kk} \right)^T * \left(A^{kl} B^{lk} F^{kk} \right) \\
+& \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k}^G  B^{lk} F^{kt} )
+ \left(V^k B^{kk} -  V^k L^{kk} \right)^T * \left(A^{kl} X^{l}  \right)\\
+& \left(V^l B^{lk} \right)^T * F^{kl}
+ \left(V^l B^{lk} \right)^T *  \left(A^{kl} L^{ll} F^{ll} \right)
+ \left(V^l B^{lk} \right)^T \\
*&  \left(A^{kl} L^{ll} E^{l*} \right) + (\sum_{t \neq k,l}^G  V^{t} B^{tk} )^{T} * F^{kl}
+ (\sum_{t \neq k,l}^G  V^{t} B^{tk} )^{T}\\
*&   \left(A^{kl} L^{ll} F^{ll} \right) + (\sum_{t \neq k,l}^G  V^{t} B^{tk} )^{T} *  \left(A^{kl} L^{ll} E^{l*} \right) ,
\end{split}
\end{align}
where $F$ is final demand, and $L$ refers to the domestic Leontief inverse as opposed to the global inverse $B$. $X$ is output while \textit{T} indicates a matrix transpose operation.

@@

\begin{itemize}
\item Domestic value added absorbed abroad (\textit{VAX\_G}, T1-5)
\begin{itemize}
\item Domestic value added in final exports (\textit{DVA\_FIN}, T1)
\item Domestic value added in intermediate exports (\textit{DVA\_INT}, T2-5)
\begin{itemize}
\item Domestic value added in intermediate exports absorbed by direct importers (\textit{DVA\_INT}, T2)
\item Domestic value added in intermediate exports re-exported to third countries (\textit{DVA\_INTrex}, T3-5)
\begin{itemize}
\item Domestic value added in intermediate exports re-exported to third countries as intermediate goods to produce domestic final goods (\textit{DVA\_INTrexI1}, T3)
\item Domestic value added in intermediate exports re-exported to third countries as  final goods (\textit{DVA\_INTrexF}, T4)
\item Domestic value added in intermediate exports re-exported to third countries as intermediate goods to produce exports (\textit{DVA\_INTrexI2}, T5)
\end{itemize}
\end{itemize}
\end{itemize}
\item Domestic value added returning home (\textit{RDV}, T6-8)
\begin{itemize}
\item Domestic value added returning home as final goods (\textit{RDV\_FIN}, T6)
\item Domestic value added returning home as final goods through third countries (\textit{RDV\_FIN2}, T7)
\item Domestic value added returning home as intermediate goods (\textit{RDV\_INT}, T8)
\end{itemize}
\item Foreign value added (\textit{FVA}, T11-12/14-15 )
\begin{itemize}
\item Foreign value added in final good exports (\textit{FVA\_FIN}, T11/14)
\begin{itemize}
\item Foreign value added in final good exports sourced from direct importer (\textit{MVA\_FIN}, T11)
\item Foreign value added in final good exports sourced from other countries (\textit{OVA\_FIN}, T14)
\end{itemize}
\item Foreign value added in intermediate good exports (\textit{FVA\_INT}, T12/15)
\begin{itemize}
\item Foreign value added in intermediate good exports sourced from direct importer (\textit{MVA\_INT}, T12)
\item Foreign value added in intermediate good exports sourced from other countries(\textit{OVA\_INT}, T15)
\end{itemize}
\end{itemize}
\item Pure double counting (\textit{PDC}, T9-10/13/16)
\begin{itemize}
\item Pure double counting from domestic source (\textit{DDC}, T9-10)
\begin{itemize}
\item Due to final goods exports production (\textit{DDF}, T9)
\item Due to intermediate goods exports production (\textit{DDI}, T10)
\end{itemize}
\item Pure double counting from foreign source (\textit{FDC}, T13/16)
\begin{itemize}
\item Due to direct importer exports production (\textit{FDF}, T13)
\item Due to other countries' exports production (\textit{FDI}, T16)
\end{itemize}
\end{itemize}
\end{itemize}

For the analysis, we use $dva\_fin$, $fva\_fin$, $rdv$, $pdc$ and the two aggregate measures $dva_inter$ combining $dva_int$ and $ddc$ as well as $fva_inter$ combining $fva_int$ and $fdc$. This collapses the indicator to a intuitive and manageable amount.

The advantage of such a detailed decomposition is that these new indicators can inform us on how countries integrate into GVCs while the basic Leontief decomposition mainly informs us on the intensity of integration. High amounts of foreign value added in final goods exports are, for instance, suggestive of a specialisation in downstream tasks that add little value to a good, such as assembly. High amounts of domestic value added in intermediate exports, on the other hand, are evidence of a more upstream specialisation in tasks that add a lot of value, such as business services. By tracking these two variables over time we can then see which countries have succeeded in moving up the value chain. We will explain the indicators in more detail in combination with the decomposition results to facilitate the understanding.

Finally, it is necessary to point out that the high resolution of the WWZ decomposition does not mean that the Leontief decomposition does not contain valuable information at all. In fact, we exploit the decomposition of exports into source industry and source country by calculating variants of the standard indicators based on different characteristics. In particular, we will assess the integration of low- and middle-income countries into GVCs by computing the amount of value added that they supply for total GVC trade.

@@

In this section we reassess with our extensive OECD ICIO dataset stylised facts on GVC integration that are typically based on smaller samples. We start by examining the development of our most basic measure of GVC integration, namely the amount of foreign value added in exports labeled by \citet{ribajalo13} as $i2e$.\footnote{Note that at the aggregate level forward ($e2r$) and backward ($i2e$) linkages are identical and thus we only look at one of the two measures.} It captures backward linkages into value chains and shows the well-known increase in GVC integration from 1995 to 2011. As illustrated in Figure \ref{fig:i2e_t}, the value of $i2e$ has grown by approximately 350\% and by 35\% as a share of total exports from around 17\% to over 23\%. Thus, countries rely for their export production increasingly on inputs produced abroad. The numbers are in line with similar findings by \citet{rojoguno12b} but their sample ends in 2009. It is then interesting to see that after the slump during the financial crises in 2009, GVCs have quickly recovered and already have started to exceed their pre-crisis levels in 2011.

Another way to examine the expansion of GVCs from 1995 to 2011 is to look at their length instead of their trade volume. WWZ propose to use the amount of double counted trade, $pdc$, as a proxy for GVC length since its value goes up the more back-and-forth trade occurs, which is equivalent to an increase in the number of production stages. They show that its value has increased for 40 selected countries. In Figure \ref{fig:pdc_t}, we observe in our larger sample similarly that $pdc$ as a share of total exports has increased over the examined period by 73\% and thus more than $i2e$. Therefore, GVCs do not only channel more trade but also have become longer over time.

@@

Turning from the development over time to sectoral differences in GVC integration, Figure \ref{fig:i2e_s} reveals in line with \citet{rojoguno12a} that the sectors exhibiting the highest degree of international fragmentation in terms of $i2e$ shares are heavy manufactures such as motor vehicles (MTR), other transport equipment (TRQ) and the metal industry (MET) as well as computers and electronics (CEQ and ELQ). In particular, the transport equipment and electronics industry are strongly engaged in GVCs having highly international production networks. For instance, Apple's iPhone contains inputs from 9 to 10 countries while the Boeing 787 production spans more than 5 countries. The sectors can be characterised as being close to final demand and producing complex differentiated goods. These characteristics can thus explain differences in GVC integration.

The bottom 6 industries in terms of $i2e$ shares are primary and services sectors such as agriculture (AGR), mining (MIN), R\&D and business services (BZS), or wholesale and retail trade (WRT). These sectors are typically located upstream in the supply chain far from final demand and have high value added to output ratios.

\begin{figure}
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{industries_i2e_top6.png}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{industries_i2e_bottom6.png}
\end{minipage}
\caption{Sectoral $i2e$ shares - Top and bottom 6.}
\label{fig:i2e_s}
\end{figure}

Naturally then, things are reversed when we look at the corresponding forward linkage GVC measure, $e2r$. It captures the amount of domestic value added in foreign exports and thus quantifies how important domestic industries are for foreign export production. Here, Figure \ref{fig:e2r_s} depicts that this indicator is dominated by the same upstream industries that are at the bottom of the $i2e$ ranking such as mining or business and telecommunication services (PTL). This shows that these industries are also strongly engaged in GVCs but their participation is of a different type. They primarily supply important inputs but do not serve final demand.

With respect to services, it is also indicative of the servicification of manufacturing as described by \citet{ribaetal15}. This means that an increasing share of manufacturing gross exports is actually value added generated in services sectors and then embedded in the intermediate goods exports of manufacturers. This importance of services sectors for exports cannot be seen from standard gross trade statistics and thus constitutes a major advantage of trade in value added measures.

It is also indicative of a growing internationalisation of services. More and more, services are being offshored and sourced from abroad. In that respect, it is also interesting to note that despite the low absolute $i2e$ shares, it is in services where a lot of the growth in $i2e$ has taken place. Five out of the six sectors with the highest growth in $i2e$ shares are services sectors.

\begin{figure}
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{industries_e2r_top6.png}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{industries_e2r_bottom6.png}
\end{minipage}
\caption{Sectoral $e2r$ shares - Top and bottom 6.}
\label{fig:e2r_s}
\end{figure}

Finally, when we turn to differences in GVC integration by country we can confirm the findings by \citet{ribajalo13}. Figure \ref{fig:i2e_k} shows that small countries close to the major GVC hubs in Asia, Europe, and North America have the highest average $i2e$ shares. Examples include Malaysia and Slovakia. Countries specialised in the primary sector or assembly on the other hand have very low values. Correspondingly, Latin American countries with their focus on agriculture and mining have very weak backward linkages into GVCs. However, the development over time shows that some of the countries with the relatively low GVC integration have begun to catch up. For instance, Argentina, India and Turkey are in the top 6 when it comes to the growth of $i2e$ shares from 1995 to 2011.

Driven by the sectoral statistics, we then find again that for $e2r$ the picture is reversed with raw material exporters on top. If we abstract from these countries we find technologically advanced countries such as Switzerland and the main GVC hubs Japan, USA, and Germany to exhibit strong forward linkages into GVCs. In particular low and middle-income countries without raw materials such as Cambodia, Mexico, or Turkey have in contrast very weak linkages and have not been able to strengthen them significantly between 1995 and 2011.\footnote{The full set of results for $i2e$ and $e2r$ by country, and sector can be found in the Appendix. Since the results of WWZ decomposition are much more detailed the results not presented here are only available from the authors upon request.}

\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{country_i2e_top_6.png}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{countries_i2e_bottom_6.png}
\end{minipage}
\caption{Countries' $i2e$ shares - Top and bottom 6.}
\label{fig:i2e_k}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{countries_e2r_top_6.png}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[width=\textwidth]{countries_e2r_bottom_6.png}
\end{minipage}
\caption{Countries' $e2r$ shares - Top and bottom 6.}
\label{fig:e2r_k}
\end{figure}

@@

The central advantage of our application is that we have new indicators for a new set of countries. This means we can not only confirm previous findings with a more representative sample but also provide several new insights. In particular, the OECD ICIO database extends the available list of countries in ICIOs by the following 21 regions: Argentina, Brunei Darussalam, Cambodia, Chile, Colombia, Costa Rica, Croatia, Hong Kong, Iceland, Israel, Malaysia, Norway, New Zealand, Philippines, Saudi Arabia, Singapore, Thailand, Tunisia, Vietnam, South Africa, and Switzerland. This means that in particular the coverage of low and middle income countries has increased considerably which allows us to analyse the GVC integration of developing economies in a much more detailed fashion than before.

\subsection{General trends in the GVC participation of developing economies}

Regarding the integration of low- and middle-income countries, \citet{rojoguno12a} have observed that per capita income is only a weak predictor for GVC integration due to the heterogeneity of economies in terms of size, industrial structure and location. In Table \ref{tab:gvc} we see that the average integration measured by either $i2e$ or $e2r$ does not vary strongly between income groups defined by the World Bank classification at the beginning of the sample period in 1995.\footnote{Note that in this section indicators are based only on manufacturing and services sectors to avoid spurious results stemming from primary sectors that are for technological reasons less integrated into GVCs.}
High-income economies have slightly stronger forward linkages but lower backward linkages which implies that their exports contain more domestic value added. Developing economies could thus try to upgrade their GVC integration by increasing domestic content in exports.
